---
title: "Network analysis"
author: "Océane"
date: "2022-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.height = 7)
library(tidyverse)
library(DIANE)
library(igraph)
library(ggraph)
library(RColorBrewer)
load(file = "../Data/network_N_CO2.RData")
```


# Context

This files loads the results from GRN inference (as given by `Transcriptomic_analysis/GRN_inference.R`), and provides various analyses of its output.


# Degree distribution of the GRN

```{r, fig.width = 5, fig.height = 4}
draw_network_degrees <- function (nodes, graph)
{
  targets <- nodes[nodes$gene_type == "Target Gene", "id"]
  TFs <- nodes[nodes$gene_type == "Regulator" | nodes$gene_type ==
                 "Grouped Regulators", "id"]
  degree_in_targets <-
    degree(graph, mode = "in", v = targets)
  degree_in_tfs <- degree(graph, mode = "in", v = TFs)
  degree_out_tfs <- degree(graph, mode = "out", v = TFs)
  betweenness <- betweenness(graph, weights = NA, v = TFs)
  deg_targ = data.frame(degree_in_targets)
  Node_nw_st <-
    data.frame(degree_in_tfs, degree_out_tfs, betweenness)
  deg_in_targ <-
    ggplot(data = deg_targ, aes(x = degree_in_targets)) +
    geom_histogram(fill = "#69b3a2",
                   alpha = 0.7) + ggtitle("In-Degree distribution of target genes") + theme_pubr() +
    theme(
      plot.title = element_text(size = 16),
      axis.text.x = element_text(size = 15),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  deg_in_tfs <-
    ggplot(data = Node_nw_st, aes(x = degree_in_tfs)) + theme_pubr() +
    geom_histogram(fill = "#69b3a2",
                   alpha = 0.7) + ggtitle("In-Degree distribution of regulators") +
    theme(
      plot.title = element_text(size = 16),
      axis.text.x = element_text(size = 15),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  deg_out_tfs <-
    ggplot(data = Node_nw_st, aes(x = degree_out_tfs)) + theme_pubr() +
    geom_histogram(fill = "#69b322",
                   alpha = 0.7) + ggtitle("Out-Degree distribution of regulators") +
    theme(
      plot.title = element_text(size = 16),
      axis.text.x = element_text(size = 15),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  bet <-
    ggplot(data = Node_nw_st, aes(x = betweenness)) + theme_pubr() +
    geom_histogram(fill = "#E69F00",
                   alpha = 0.7) + ggtitle("Betweeness distribution of regulators") +
    theme(
      plot.title = element_text(size = 16),
      axis.text.x = element_text(size = 15),
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )
  return(deg_in_targ + deg_in_tfs + deg_out_tfs + bet)
}

draw_network_degrees(network_N_CO2$network_data$nodes, network_N_CO2$network)
```

# Ranking the candidate genes

The final ranking is made of the top 12 regulators in terms of overall degree.

```{r}
edges <- network_N_CO2$network_data$edges
nodes <- network_N_CO2$network_data$nodes
table(nodes$group)

# annotation of candidate regulators in the network
labs <- toupper(
  c(
    "myb93",
    "tcp8",
    "wox11",
    "wrky59",
    "edf3",
    "trfl6",
    "zf2",
    "agl14",
    "myb85",
    "myb15"
  )
)
annot <- gene_annotations$`Arabidopsis thaliana`
annot$gene <- rownames(annot)
candidates <- annot[annot$label %in% labs |
                      annot$gene %in%
                      c("AT5G43175", "AT2G29060"),
                    c("label", "gene")]
nodes$candidate <-
  ifelse(str_detect(nodes$id, paste(candidates$gene, collapse = '|')),
         "Candidate", "")
# annotation of nitrate genes in the network
n_genes <- read.csv(file = "../Data/N-related-genes-in-network.csv", sep = ';', h = T)
nodes$nitrate_gene <- ifelse(
  str_detect(nodes$id,paste(n_genes$AGI, collapse = '|')), 
  "N-related gene", "")


ranking <- nodes[order(-nodes$degree),]
ranking$rank <- 1:nrow(ranking)
final_ranking <- ranking[1:12,c("label", "community_sbm", "degree", "rank")]
colnames(final_ranking)[2] <- "community"
final_ranking["AT5G43175", 'label'] <- "AT5G43175 (bHLH)"
final_ranking$rank[11:12] <- c(10,10);final_ranking
```
# Network visualisation


```{r}
# removing little disconnected subnetworks for visualisation
graph <- graph_from_data_frame(edges)
components <- clusters(graph, mode="weak")
biggest_cluster_id <- which.max(components$csize)
vert_ids <- V(graph)[components$membership == biggest_cluster_id]
graph <- induced_subgraph(graph, vert_ids)

# creation of layout object
layout <- ggraph::create_layout(graph, layout = "stress")
layout[c("label", "degree", "community_sbm", "gene_type", "candidate", "nitrate_gene")] <- 
  nodes[match(layout$name, nodes$id), 
                 c("label", "degree", "community_sbm", "gene_type", "candidate", "nitrate_gene")]
layout$regulator <- ifelse(str_detect(layout$gene_type, "egul"), "Regulator", "Target")
layout$regulator <- relevel(factor(layout$regulator), ref = "Target")
layout$size <- ifelse(str_detect(layout$gene_type, "ouped"), "Grouped regulators", "other")
layout$size <- relevel(factor(layout$size), ref = "other")

layout$gene_type <- fct_relevel(factor(layout$gene_type), 
                                c("Regulator", "Grouped Regulators", "Target Gene"))
layout$candidate_label <- ifelse(layout$candidate == "Candidate", layout$label, "")
layout$nitrate_label <- ifelse(layout$nitrate_gene == "N-related gene", layout$label, "")

```

## Network view with candidates highlighted

```{r, fig.width = 4, fig.height = 4}
cands <- ggraph(layout) + 
  geom_edge_link(color = "grey") + 
  geom_node_point(aes(shape = regulator, size = candidate,
                      fill = candidate , alpha = candidate), stroke = 0.3, 
                  color = "black", show.legend = F) +
  scale_shape_manual(values=c(21,22)) + 
  scale_size_manual(values = c(1.5, 5)) + 
  scale_alpha_manual(values = c(0.3, 1))+ 
  scale_fill_manual(values = c("lightgrey", '#cd5b45'))+
  geom_node_label(aes(label = candidate_label, alpha = candidate, color = community_sbm),
                  color = '#cd5b45', size = 3.5, nudge_y = 0,
                  label.padding = 0.1, repel = TRUE, 
                  max.overlaps = 100, show.legend = F) + 
  theme_void();cands
```

## Network view with nitrate-related genes highlighted

```{r, fig.width = 4, fig.height = 4}
n_related <- ggraph(layout) + 
  geom_edge_link(color = "grey") + 
  geom_node_point(aes(shape = regulator, size = nitrate_gene,
                      fill = nitrate_gene , alpha = nitrate_gene), stroke = 0.3, 
                  color = "black", show.legend = F) +
  scale_shape_manual(values=c(21,22)) + 
  scale_size_manual(values = c(1.5, 4.5)) + 
  scale_alpha_manual(values = c(0.3, 1))+ 
  scale_fill_manual(values = c("lightgrey", '#1f9f7b'))+
  geom_node_label(aes(label = nitrate_label, alpha = nitrate_gene, color = community_sbm),
                  color = '#1f9f7b', size = 3.5, nudge_y = 0.1,
                  label.padding = 0.1, repel = TRUE, 
                  max.overlaps = 200, show.legend = F) + 
  theme_void();n_related
```

## Network view with topological clusters

Modules given by SBM clustering are shown below.


```{r, fig.width = 4, fig.height = 4}
nb.cols <- max(layout$community_sbm)
mycolors <- colorRampPalette(brewer.pal(11, "Paired"))(nb.cols)

sbm_comm <- ggraph(layout) +
  geom_edge_link(color = "grey") + 
  geom_node_point(aes(shape = regulator, size = gene_type,
                      fill = factor(community_sbm)), stroke = 0.25, 
                  color = "black", show.legend = T) +
  scale_shape_manual(values=c(21,22), guide = 'none') + 
  scale_size_manual(values = c(2,3,1.5)*1.8, guide = 'none') + theme_void()+
  scale_fill_manual(name = "Community\nmembership", 
                    values = mycolors) + 
  guides(fill = guide_legend(override.aes = list(shape = 21, size = 5))) + 
  theme(legend.position = "bottom", 
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)); sbm_comm
```


# Network validation against connecTF

```{r, fig.width = 4, fig.height = 3}
source("evaluateNetwork.R")
results <- evaluate_network(edges)
draw_evaluated_network <- function (results)
{
  biggest_component_only = TRUE
  graph <- graph_from_data_frame(results$edges)
  if (biggest_component_only) {
    components <- clusters(graph, mode = "weak")
    biggest_cluster_id <- which.max(components$csize)
    vert_ids <- V(graph)[components$membership ==
                           biggest_cluster_id]
    graph <- induced_subgraph(graph, vert_ids)
  }
  layout <- ggraph::create_layout(graph, layout = "stress")
  layout$gene_type <-
    ifelse(layout$name %in% edges$from, "Regulator",
           "Target")
  layout$gene_type <- ifelse(
    stringr::str_detect(layout$name,
                        "mean_"),
    "Grouped regulators",
    layout$gene_type
  )
  layout$regulator <-
    relevel(factor(layout$gene_type), ref = "Target")
  if ("Grouped regulators" %in% layout$gene_type)
    layout$gene_type <-
    forcats::fct_relevel(factor(layout$gene_type),
                         c("Regulator", "Grouped regulators",
                           "Target"))
  else
    layout$gene_type <- forcats::fct_relevel(factor(layout$gene_type),
                                             c("Regulator", "Target"))
  layout$regulator <- ifelse(stringr::str_detect(layout$gene_type,
                                                 "egulator"),
                             "Regulator",
                             "Target")
  layout$regulator <-
    relevel(factor(layout$regulator), ref = "Target")
  colors <-
    c(
      `Not supported` = "#393939",
      `TF not studied` = "grey",
      CHIPSeq = "#0C7C59",
      DAPSeq = "#aceca1",
      TARGET = "#FAC05E",
      `CHIPSeq+DAPSeq` = "red",
      `DAPSeq+TARGET` = "#749bcd",
      `TARGET+DAPSeq` = "#749bcd",
      `CHIPSeq+DAPSeq+TARGET` = "green"
    )
  val <-
    ggraph::ggraph(layout) + ggraph::geom_edge_link(
      aes(
        alpha = !stringr::str_detect(type,
                                     "TF not studied"),
        col = type
      ),
      width = 0.71,
      show.legend = TRUE
    ) +
    ggraph::geom_node_point(
      aes(
        shape = regulator,
        size = gene_type,
        fill = gene_type
      ),
      stroke = 0.235,
      color = "black",
      show.legend = F,
      
    ) + ggraph::theme_graph() +
    scale_shape_manual(values = c(21, 22), guide = "none") +
    scale_size_manual(values = c(2, 3, 1.5) * 0.95,
                      guide = "none") +
    ggraph::scale_edge_color_manual(name = "Nature of the \nsupporting knwon\ninteraction",
                                    values = colors) +
    labs(
      title = "Network edges colored according to their experimental support",
      subtitle = paste(
        round(results$tpr * 100, 2),
        "% of the edges (with validation information available) are supported"
      )
    ) +
    ggraph::scale_edge_alpha_manual(values = c(0.51, 1),
                                    guide = "none")
  if ("Grouped regulators" %in% layout$gene_type)
    val <- val + scale_fill_manual(values = c("#68C872",
                                              "darkgreen", "grey"),
                                   guide = "none")
  else
    val <- val + scale_fill_manual(values = c("#68C872",
                                              "grey"), guide = "none")
  
  return(val)
}
draw_evaluated_network(results)
```

